Web Processing - Standardised GIS Analyses for Cable Route Planning

### 2023-03-08 paper/discussion.tex
Updated Discussion & Conclussion
readability test for the first two sections

Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100~m!
No problem because buffering is included in the cost raster.
Not algorithmic speed-ups as pre-compilation, JIT, vector units, multi-core, weight range

>>>>

The fact that the differences in the costs between the paths from the all touched false to the all touched true rasters stay almost constant over the different resolution is probably due to the fact that, the costs per resolution stays in narrow range of costs.
This effect of the similar aggregated costs per resolution could also be seen in the testing paths, even when the paths changed strongly with the change of resolution or algorithm applied.
Which might be a hint for an evenly spatial distribution of the costs.

The all touched true cost raster show every detail, but the sampling with all touched true increases the size of the features.
The fact that the aggregated costs per resolution for all touched true rasters overestimates the costs when computing the path from a low resolution raster, might be due to the fact, that the values of the costs are un-evenly distributed, but that the high costs are much more frequent differ and the costs are exponential scaled exponentially.

### 2023-03-08 paper/methods.tex
Updated Discussion & Conclussion
readability test for the first two sections

PyWPS~\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
The initial implementation of the least cost path algorithm is based on the implementation for the QGIS-Plugin
'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.

The different layers from the different entities are optionally filtered, buffered and then rasterized.
Filtering the layers of the files for special attributes enables to further differentiate.
For examples makes it possible to differentiate between heath and uncultivated land.
Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the
correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
The costs of all layers of is aggregated with the maximum function.
Thus, an area covered by multiple layers is uniformly used with the highest costs.
For every area in the study area, that is not covered by any layer, is given the default cost.
The costs have been grouped into different levels~(see table~\ref{tab:1}) starting from \textit{preferential} areas with
very low costs, via \textit{no restriction}s, which is the default, used when no other layers are covering the local area,
to \textit{restricted}, \textit{strongly restricted} and \textit{prohibited} areas with high costs.
These higher costs resemble the degree how much a local area should be avoided, while routing the path.
The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels
the algorithm is willing to go, for avoiding an area of high costs.
Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
the weight that resembles the costs for these types of areas, has to be especially high. \\
The rasterization transforms a vector into a raster.
The rasterization can be executed in to different ways.
In both ways, the rasterization can be imaged, as the old vector is layered over the new raster grid with the new
given resolution and the new affine transformation and the coordinate reference system of the vector.
Both rasterization techniques differ in how the pixels are selected, that describe, that a geo-object overlays the pixel.
The pixels can either be described as overlaid with a geo-object, when either the center of that pixel is overlaid
by the geo-object, or any part of the pixel is overlaid.
Hence, the version with any part overlaid is called all touched True.
The version where a overlaid pixel center is require is thus called all touched False.
All touched False is considered the default (see figure~\ref{fig:alltouched}).
\begin{figure}[!ht]
	\centering

@@

\begin{table}[h!]
	\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
	\label{tab:1}
	\centering
	\begin{tabular}{ l  r  l }
		Cost Level 			& Cost 					& Example\\
		\hline
		Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
		strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
		Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
		No Restriction 		& 0.5					& Default\\
		Preferential 		& 0.1					& \makecell[lt]{Power Grid,\\ Motorway and Railway Buffers}\\
	\end{tabular}
\end{table}
The completed list of layers and the processing applied to them, can be found in \textbf{Supplement S1}.

The whole stack of the cost raster, via the aggregated costs to the resulting least cost path can is shown with an example for a starting cost raster of 50~m resolution and a rasterization with all touched False (see \ref{fig:costs2path}).
Because of the early stopping strategy used in the aggregation sub-paths with high costs have not been fully computed.
Instead, the execution stops early after reaching the end point.
Because the end point ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} Area.
Therefore, areas even farther away from the starting point have been explored first.

@@

\todo{Were to put the following part?}
The aggregation is implemented in a priority queue. 
Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.
The aggregation is implemented with early stopping. 
After the last end point has been has got an aggregated cost the aggregation stops and the back tracking starts.

>>>>

PyWPS~\cite{noauthor_welcome_2016} is used to provide the least cost path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sending the cost raster and receiving the resulting least cost path.
The initial implementation of the least cost path algorithm is based on the implementation for the QGIS-Plugin
'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.

The different layers from the different entities are optionally filtered, buffered and then rasterized.
Filtering the layers of the files for special attributes enables to further differentiate.
For examples makes it possible to differentiate between heath and uncultivated land in the land coverage.
Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the
correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
The costs of all layers is aggregated with the maximum function.
Thus, an area covered by multiple layers is uniformly used with the highest costs.
Any place in the study area, that is not covered by any layer and thus does not have a weight yet, is given the default cost.

The costs have been grouped into different levels~(see table~\ref{tab:1}) starting from \textit{preferential} areas with
very low costs, via \textit{no restriction}s, which is the default, used when no other layers are covering the local area,
to \textit{restricted}, \textit{strongly restricted} and \textit{prohibited} areas with high costs.
These higher costs resemble the degree how much a place of this cost should be avoided, while routing the path.
The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels
the algorithm is willing to go, for avoiding an area of this high costs.
Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
the weight that resembles the costs for these types of areas, has to be especially high.\\

All these layers are provided as vectors.
While the least cost algorithm use raster data.
The rasterization transforms a vector into a raster.
The rasterization can be executed in two different ways.
In both ways, the rasterization can be imaged, as the old vector is layered over the new raster grid with the new
given resolution and the new affine transformation and the coordinate reference system of the vector.
Both rasterization techniques differ in the selection of the pixel, that describe the original polygon.
The pixels can be selected , if either the center of that pixel is overlaid
by the geo-object, or any part of the pixel is overlaid.
The version with any part overlaid is called all touched True.
The version where a overlaid pixel center is require is called all touched False.
All touched False is considered the default (see figure~\ref{fig:alltouched}).
\begin{figure}[!ht]
	\centering

@@

\begin{table}[h!]
	\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
	\label{tab:1}
	\centering
	\begin{tabular}{ l  r  l }
		Cost Level 			& Cost 					& Example\\
		\hline
		Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
		strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
		Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
		No Restriction 		& 0.5					& Default\\
		Preferential 		& 0.1					& \makecell[lt]{Power Grid,\\ Motorway and Railway Buffers}\\
	\end{tabular}
\end{table}

The completed list of layers and their processing, can be found in \textbf{Supplement S1}.

All three steps of the generation of the least cost path: generation of the cost raster, aggregation and back tracking is shown with an example for a starting cost raster of 50~m resolution and a rasterization with all touched False (see \ref{fig:costs2path}).

The chosen implementation applies early stopping.
Therefore, the costs for points that are not needed to try to connect to the end point are not aggregated.
After an aggregated cost for every end point has been found, the aggregation stops and the back tracking starts.
Because the end point ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} Area.
Therefore, areas even farther away from the starting point have been explored first.

@@

%% The aggregation is implemented in a priority queue. 
%% Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.

### 2023-03-08 paper/paper-CableRoute.tex
Updated Discussion & Conclussion
readability test for the first two sections

	\begin{abstract}
		The least cost paths for a power line route from a power station near the German shore line is computed to a power station further inland.
		Down sampling and clipping studied for a speed-up of the computation.
		The path from clipped cost raster often matches the original path.
		The speed-up of the clipping can not be known before hand.
	\end{abstract}

@@

	The German power system add a lot of renewable energy sources in the north, because wind energy reaches highest efficiency with high wind speeds.
	At the same time old power plants e.g. nuclear, hard coal and lignite are going out off service~\cite{eser-frey_analyzing_2012}.
	These older power plants were mainly situated in southern and central Germany. 
	At the same time , the power sink industrial and private demand do not shift northwards.
	Therefore, the renewable energies have to be transported from north to south which increases the congestion in the power grid. 
	The amount of offshore wind power, that the German energy system can use, can be greatly increased by adding new power lines~\cite{leuthold_nodal_2005}.

	But for building power lines, finding the shortest path is not sufficient.
	Additional parameters also have to be taken into consideration.
	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
	For planing the additional routes for a power grid, further aspects as legal regulations and acceptance
	by the local population have to be taken in consideration.
	Also technical aspects, as the effects on the grid stability might be further points to consider~\cite{schafer_understanding_2022}.

	Due to the increasing need for renewable energy wind off shore wind turbines are source which currently supplies 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}
.	
	Highest wind speeds are reached along the cost lines in the north of Germany.

	Other methods as the multi-criteria decision analysis take further aspects and stakeholder as experts, public and decision maker and into consideration~\cite{bertsch_participatory_2016}.

	For planning a possible route for a power line the least cost path algorithm is applied.
	The least cost path algorithm is a Dijkstra algorithm applied on a raster map.
	The vertices of the graph are the centers on neighboring pixels in al possible eight directions.
	This enables to find routes on graphs that are not predefined like road notworks.
	The weights between the edges are the local costs to travel from one pixel center to the neighboring pixel center.
	The costs might be physical costs, as the local slope, but also can be composed on other factors as the acceptance rate for the land usage of a given class.
	The least cost path algorithm consists of at least two sequential steps.
	1) The first steps is to aggregate the costs of traveling from the starting point to a given set of end points.
	This steps generates the aggregated cost raster of the total cost that are needed to travel to a point, from the cost raster.
	2) In the next step the back tracking, the route of the actual least cost path is calculated.
	For each ending point the path via the lowest cost neighbor is taken until arriving at the starting point.

@@

	Older models are e.g. are optimizing the well fare, economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.

@@

	\section{Conclusion}\label{sec:conclusion}
	When using the researched strategies, the on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
	\todo{survey: local population, relative weights per district?}

>>>>

	\begin{abstract}
		The least cost paths for a power line route from a power station is computed to a power station further inland.
		Down sampling and clipping as studied as techniques to speed-up the computation.
		The path from clipped cost raster often matches the original path.
		The speed-up of the clipping can not be known before hand.
	\end{abstract}

@@

	The German power system adds a lot of renewable energy sources in the north, because wind energy reaches highest efficiency with high wind speeds.
	At the same time old power plants e.g.\ nuclear, hard coal and lignite are going out off service~\cite{eser-frey_analyzing_2012}.
	These older power plants were mainly situated in southern and central Germany. 
	The power sink, which is industrial and private demand does not shift northwards.
	Therefore, the renewable energies have to be transported from north to south which increases the congestion in the power grid. 
	The amount of offshore wind power, that the German energy system can use, can be greatly increased by adding new power lines~\cite{leuthold_nodal_2005}.

	But for building power lines, finding the shortest path is not sufficient.
	Additional parameters also have to be taken into consideration.
	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
	For planing the additional routes for a power grid, further aspects as legal regulations and acceptance
	by the local population have to be taken in consideration.
	Also technical aspects, as the effects on the grid stability are further points take into consideration~\cite{schafer_understanding_2022}.

	Due to the increasing need for renewable energy wind offshore wind turbines supply 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.

	Other methods as the \acrfull{MCDA}  consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).

	For planning a possible route for a power line the least cost path algorithm is applied.
	The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
	The vertices of the graph are the pixel centers, that are connected to the eight neighboring pixel centers via the edges.
	This enables to find routes on graphs that are not predefined like road notworks.
	The weights of the edges are the local costs to travel from one pixel center to the neighboring pixel center.
	The costs might be physical costs, as the local slope, but also can be composed on other factors as the acceptance rate for the land usage of a given class.
	The least cost path algorithm consists of at least two sequential steps.
	1) The first step is to aggregate the costs of traveling from the starting point to a given set of end points.
	This step generates the aggregated cost raster of traveling from the starting point to any point of the cost raster.
	2) In the next step the back tracking, the route of the actual least cost path is calculated.
	For each ending point the path via the lowest cost neighbor is taken until arriving at the starting point.

	Some implementations switch the role of start and ending points, so that either many starting points and one end point, or one end point and many endpoints can be used.
	In some implementation there is a extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
	path to the starting point as integer values.

@@

	Older models are e.g.\ are optimizing the well fare, economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.

@@

	\section{Conclusion}\label{sec:conclusion}
	The cost of computation of the least cost path scales with the square of the resolution.
	The difference between the aggregated costs of the paths per resolution computed from the all touched false raster and the all touched true raster only shows a linear decrease.
	Hence, the gain in precision per compute time decreases.
	The presented methods tried to circumvent this bottleneck.
	When using the researched strategies, on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
	The paths might change very strongly even with small changes in total costs.
	Hence to develop an alternative backtracking algorithm that, that generates a set of possible paths within a corridor of costs might be a good strategy, to offer alternative routes and show variations between these routes at the same time.
	An alternative method for future methods, would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster. 
	This in turn increases the computation time again.

	In an actual search for least cost path a survey might be used as method to estimate the costs.
	When the sample size of the survey is large enough, the
	weights might take local differences relative acceptance into consideration.

	Only changes in the algorithm have been applied in the search for a speed-up.
	There a still other methods as just-in-time compilation.

### 2023-03-08 paper/results.tex
Updated Discussion & Conclussion
readability test for the first two sections

At the same time the differences in the aggregated costs stay constant.
\todo{Interpretation} When normalize the aggregated costs by the resolution.
Than On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
linear with the resolution.
On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
the resolution.
Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.

@@

The zonal stat (see table~\ref{tab:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
percentage of which costs levels are around the path.
When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
\textit{Preferential Level} and less the\textit{NoRestriction} Level.\todo{Interpretation:}
The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.

@@

For three of the four routes the least cost path computation from the clipped raster, was able to compute the expected result.
For the forth path, the
least cost path from the 5~m resolution raster was out side of the buffer around the least cost path from the 50~m resolution raster.
The speed up from the clipping of the higher resolution raster depends on the amount of pixels that, have been clipped. 
Bi-linear down sampling the of the high resolution raster to a medium resolution, did not result in any benefits compared to an original medium resolution raster.
The resolution corrected aggregated costs of the least cost path from the down sampled raster are higher, than those, from the higher 5~m resolution raster and the medium 10~m resolution raster.
In addition, is the distance, from this path to the path from the high resolution raster much higher, than the distance of path from the original 10~m resolution raster to the path from the 5~m resolution raster (see table~\ref{tab:6}).
\todo{interpretation Why, down sampling so different???}

>>>>

At the same time the differences in the aggregated costs stay almost constant.
When normalize the aggregated costs by the resolution.
On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
linear with the resolution.
On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
the resolution.
Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.

@@

The zonal stat (see table~\ref{tab:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
percentage of which costs levels are around the path.
When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
\textit{Preferential Level} and less the\textit{NoRestriction} Level.
The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.

@@

For three of the four routes the least cost path computation from the clipped raster, was able to compute the expected result.
For the forth path, the
least cost path from the 5~m resolution raster was out side of the buffer around the least cost path from the 50~m resolution raster.
The speed up from the clipping of the higher resolution raster depends on the amount of pixels that, have been clipped. 
Bi-linear down sampling the of the high resolution raster to a medium resolution, did not result in any benefits compared to an original medium resolution raster.
The resolution corrected aggregated costs of the least cost path from the down sampled raster are higher, than those, from the higher 5~m resolution raster and the medium 10~m resolution raster.
In addition, is the distance, from this path to the path from the high resolution raster much higher, than the distance of path from the original 10~m resolution raster to the path from the 5~m resolution raster (see table~\ref{tab:6}).
\todo{interpretation: Why result from down sampling so different? Don't know.}
